Investigation of layer specific BOLD signal in the human visual cortex during visual attention

### 2022-02-27 paper.tex
spellcheck

Because of temporal changes in magnetic field inhomogeneity, local topology slightly changed over the course of the entire session. For this reason, the 18 functional runs obtained for each subject were first divided into three groups of each 6 contiguous runs, and then each group was pre-processed separately. Time courses were subsequently concatenated before entering the main analyses.

@@

\subsection{Temporal analysis}
Temporal linear regression was used to compare between the experimental conditions. Regressors were created as follows. The stimuli appeared during the stimulus window on 2/3rds of trials, which were modeled with a single regressor (stimulus on). The remaining stimulus windows were also modeled with a regressor (stimulus off). In addition, attention could be directed to either the left or right visual field; these conditions were each modeled with a regressor. We so obtained four regressors for each of the conditions of interest. Thus, for a given retinotopic region of interest, the four different conditions were: stimulus, no stimulus, attended, and unattended. We used a double-gamma function, as defined by SPM (parameters: time-to-peak first gamma: 5 second, time-to-peak second gamma: 10 seconds, amplitude ratio: 2:1), to model the fMRI responses. These parameters were established based on an initial finite impulse response (FIR) analysis \citep{Friston1998} using the visual stimulus response from four pilot subjects (not included in the current study). Based on the observed fMRI response in this pilot data set, temporal or dispersion derivatives were not included into the statistical model of the main experiment. The baseline signal of each run was captured by adding a regressor column of ones for each run separately. As described above, the data were pre-processed by means of nuisance regression. This was performed by adding the nuisance regressors to the design matrix, effectively adjusting for the statistical loss in degrees of freedom as a result of nuisance regression. 
% The reference of one percent signal change was the height of a peak of a two-second-long isolated event \citep{Mumford2007}.

@@

As control analyses, we recomputed the layer-specific ANOVA in a Baysian framework. The Baysian analysis gives an indication of what model best explains the observed results with the least number of parameters, expressed as the likelihood of the models. We further recomputed the original ANOVA with two different preprocessing straties. For the first, we applied a simple vascular model that should account for the blood flow effects towards downstream layers \citet{Markuerkiaga2021}. The model describes that by approximation 20\% of the layer signal seeps through to downstream layers. This can be accounted for by means of a spatial deconvolution. For the second, we z-scored the time courses per run \citet{Lawrence2018}, prior to the main analysis.

>>>>

Because of temporal changes in magnetic field inhomogeneity, local topology slightly changed over the course of the entire session. For this reason, the 18 functional runs obtained for each subject were first divided into three groups of each 6 contiguous runs, and then each group was preprocessed separately. Time courses were subsequently concatenated before entering the main analyses.

@@

\subsection{Temporal analysis}
Temporal linear regression was used to compare between the experimental conditions. Regressors were created as follows. The stimuli appeared during the stimulus window on 2/3rds of trials, which were modeled with a single regressor (stimulus on). The remaining stimulus windows were also modeled with a regressor (stimulus off). In addition, attention could be directed to either the left or right visual field; these conditions were each modeled with a regressor. We so obtained four regressors for each of the conditions of interest. Thus, for a given retinotopic region of interest, the four different conditions were: stimulus, no stimulus, attended, and unattended. We used a double-gamma function, as defined by SPM (parameters: time-to-peak first gamma: 5 second, time-to-peak second gamma: 10 seconds, amplitude ratio: 2:1), to model the fMRI responses. These parameters were established based on an initial finite impulse response (FIR) analysis \citep{Friston1998} using the visual stimulus response from four pilot subjects (not included in the current study). Based on the observed fMRI response in this pilot data set, temporal or dispersion derivatives were not included into the statistical model of the main experiment. The baseline signal of each run was captured by adding a regressor column of ones for each run separately. As described above, the data were preprocessed by means of nuisance regression. This was performed by adding the nuisance regressors to the design matrix, effectively adjusting for the statistical loss in degrees of freedom as a result of nuisance regression. 
% The reference of one percent signal change was the height of a peak of a two-second-long isolated event \citep{Mumford2007}.

@@

As control analyses, we recomputed the layer-specific ANOVA in a Bayesian framework. The Bayesian analysis gives an indication of what model best explains the observed results with the least number of parameters, expressed as the likelihood of the models. We further recomputed the original ANOVA with two different preprocessing strategies. For the first, we applied a simple vascular model that should account for the blood flow effects towards downstream layers \citet{Markuerkiaga2021}. The model describes that by approximation 20\% of the layer signal seeps through to downstream layers. This can be accounted for by means of a spatial deconvolution. For the second, we z-scored the time courses per run \citet{Lawrence2018}, prior to the main analysis.
